\SW: Crowd-sourcing the Discovery of Gravitational Lenses

### 2014-06-25 doc/sw-system.tex
Wrote up information theory from Taipei

We define the following properties of the crowd, as characterised by their
agents, and plot the distributions of their logarithms in
\Fref{fig:crowd:cornerplot}. 
%
\begin{description}
%
\item{\noindent\bf ``Effort:''} The number of test images, $\effort$, classified by a
volunteer. In stage 1, the mean effort per agent was 263; in the shorter stage
2 it was 81.
%
\item{\noindent\bf ``Experience:''} The number of training images, $\experience$,
classified by a volunteer. In stage 1, the mean experience per agent was 29;
in stage 2 (where the training image fequency was set higher) it was 34.
%
\item{\noindent\bf ``Skill:''} The expectation value of the information gain, $\skill$
should  the next subject classified have lens probability 0.5
\Aref{appendix:swap},  in bits. Random classifiers have $\skill = 0.0$,
perfect classifiers have $\skill = 1.0$. All agents start with $\skill = 0.0$;
this increases as training subjects are classified, and the agent's estimates
of its confusion matrix elements improve. In stage 1, the mean skill per agent
was 0.04 bits; in  stage 2 it was 0.05.
%
\item{\noindent\bf ``Contribution:''} The integrated skill over a volunteer's test
subject classification history, and represents the total  contribution to the
project that volunteer (see the appendix for more discussion of this
quantity). In stage 1, the mean contribution per agent was 34.9 bits; in 
stage 2 it was 33.5.
%
\item{\noindent\bf ``Information:''} The total information $\information$ generated by
the  agent during the volunteer's classification activity. This quantity
depends on the value of each subject's lens probability when that subject was 
presented to the volunteer (\Aref{appendix:swap}), and so there is an element
of luck involved with this quantity: if you never see a high probability
subject, it's hard to generate a large amount of information. You make your
own luck by classifying more subjects.
%
\end{description}

@@

\Tref{tab:crowd:contributions} shows the total effort, contribution, skill and
information generated in both stage 1 and 2 of the CFHTLS project, with the
total numbers of agents and subjects for comparison. These numbers allow us to
quantify the efficiency of the system. The contribution per classification is
defined in terms of a hypothetical subject with lens probability of 0.5; one
bit of information is needed to update such a subject's lens probability to
either zero or one. This means that a maximally complete classification stage
would yield a total contribution (summed over all agents) equal to the number
of subjects. The ratio of this hypthetical optimum to the actual total
contribution is therefore a measure of the stage's inefficiency. We find our
inefficiency (by dividing column 2 by column 3) to be 33\% and 17\% in stage 1
and 2 respectively. In stage 1, this inefficiency is due to the daily
processing: we were not able to retire subjects fast enough, and so they,
remained in the system, being over-classified. Only 3705745 classifications
were needed to retire all the subjects: the ratio of this to the total number
made is 34\%. (The remaining 1\% is due to not all subjects being classified
to 1 or 0 probability.) At stage 2, we did not retire any subjects at all; the
inefficiency in this case was by design, to give everyone a chance to
appreciate what they had found together!

@@

The total information generated during the survey is harder to interpret,
because it's exactly the high probability candidates that we leave in the
system that give rise to most of the information generated. What we can do is
divide the total information generated by the amount of information it takes
to classify a \sw subject all the way to the detection threshold (lens
probability 0.95), divide by this number, and then multiply by the survey
inefficiency to get a very rough estimate for the effective number of
detections corresponding to the crowd's contribution. With a prior probability
of $2\times10^{-4}$, it takes 11.6 bits to make a detection; the effective
number of detections contributed by the stage 1 and stage 2 volunteers is then
2670 and 24 respectively. These figures are close to the numbers of detections
given in column 7 of \Tref{tab:crowd:contributions}. 
The uncertainty in the interpretation of the
information generated provides further justification for our focus on the
expected information gain as a measure of volunteer contribution.

@@

The full confusion matrix of the $k^{\rm th}$ volunteer's agent is therefore:
\be
  \mathcal{M}^k = 
  \begin{bmatrix}
    \pr(\saidLENS|\NOT,T_k) & \pr(\saidLENS|\LENS,T_k) \\
    \pr(\saidNOT |\NOT,T_k) & \pr(\saidNOT |\LENS,T_k)
  \end{bmatrix}.
\ee
Note that these probabilities are normalized, such that
$\pr(\saidNOT |\NOT) = 1 - \pr(\saidLENS|\NOT)$.

@@

Suppose we have a volunteer who is always right about the true
nature of a training subject. 
Their agent's confusion matrix would be
\be
  \mathcal{M}^{\rm perfect} = 
  \begin{bmatrix}
    0.0 & 1.0 \\
    1.0 & 0.0
  \end{bmatrix}.
\ee
On being given a fresh subject that actually is a $\LENS$, this hypothetical
volunteer would submit $C = \saidLENS$.  Their agent would then calculate the
posterior probability for the subject being a $LENS$ to be
\begin{align}
  \pr(\LENS|\saidLENS,T_k) &= \frac{1.0 \cdot p_0}
           {\left[ 1.0\cdot p_0 + 0.0\cdot(1 - p_0) \right]}
   &= 1.0,
\end{align}
as we might expect for such a {\it perfect} classifier.  Meanwhile, a
hypothetical volunteer who (for some reason) wilfully always submits the wrong
classification would have an agent with the column-swapped confusion matrix
\be
  \mathcal{M}^{\rm obtuse} = 
  \begin{bmatrix}
    1.0 & 0.0 \\
    0.0 & 1.0
  \end{bmatrix},
\ee
and would submit $C = \saidNOT$ for this subject. However, such a volunteer
would nevertheless be submitting useful information, since given the above
confusion matrix, their agent would calculate
\begin{align}
  \pr(\LENS|\saidNOT,T_k) &= \frac{1.0 \cdot p_0}
           {\left[ 1.0\cdot p_0 + 0.0\cdot(1 - p_0) \right]}
   &= 1.0.
\end{align}
{\it Obtuse} classifiers turn out to be as helpful as {\it perfect} ones.

\subsection{Information Contribution}
\label{appendix:swap:examples}

The information likely to be contributed by each agent for a given subject
can be estimated before the next classification of that subject is made, 
just from its confusion matrix.
The Shannon entropy generated by a classifier upon performing a
classification is
\be
  \langle S_k \rangle =  
   -P_{\rm right}\cdot\log_2{P_{\rm right}} -  P_{\rm wrong}\cdot\log_2P_{\rm wrong}\,,
\ee
where $P_{\rm right}$ and $P_{\rm wrong}$ are the averages of the diagonal and
the off-diagonal elements of the confusion matrix, respectively, and $\langle
S_k \rangle$ is measured in ``bits.'' These averages represent the probability
of a classifier to get a classification right or wrong, respectively. We define
the information contributed by a classifier as
\begin{equation}
I_k = 1 - S_k\,.
  \label{eq:app:info}
\end{equation}
\Eref{eq:app:info} gives the required result, that both the hypothetical {\it
perfect} and {\it obtuse} classifiers contribute 1 bit of information each, per
classification. Classifiers whose agent's confusion matrix is such that $P_{\rm
right}=P_{\rm wrong}=0.5$, contribute zero bits of information. Such users
identify a lens correctly with the same probability as they misclassify a dud
image to contain a lens, and thus their classification is of no value.

We conservatively initialise all the elements of the agents' confusion matrices
to be 0.5, that of a random classifier. This makes no allowance for volunteers
that actually do have previous experience of what gravitational lenses look
like, but should help prevent large numbers of false positives being assigned
high probability. Plotting  $\langle I_k \rangle$ as a function of time will,
to some extent, illustrate the learning process undergone by the $k^{\rm th}$
volunteer-agent partnership.

@@

@@

The confusion matrix obtained from the application of Equation
(\ref{eq:app:fraction}) has some inherent noise which reduces as the number of
training subjects classified by the agent's volunteer increases. For
simplicity, the discussion thus far assumed the case when the confusion matrix
is known perfectly; in practice, we allow for uncertainty in the agent
confusion matrices by averaging over a small number of samples drawn from
Binomial distributions characterised by the matrix elements
$\pr(C_k|\LENS,T_k)$ and  $\pr(C_k|\NOT,T_k)$. The associated standard
deviation in the estimated subject probability provides an error bar for this
quantity.

>>>>

We define the following properties of the crowd, as characterised by their
agents, and plot the distributions of their logarithms in
\Fref{fig:crowd:cornerplot}. 
%
\begin{description}
%
\item{\noindent\bf ``Effort:''} The number of test images, $\effort$, classified by a
volunteer. In stage 1, the mean effort per agent was 263; in the shorter stage
2 it was 81.
%
\item{\noindent\bf ``Experience:''} The number of training images, $\experience$,
classified by a volunteer. In stage 1, the mean experience per agent was 29;
in stage 2 (where the training image fequency was set higher) it was 34.
%
\item{\noindent\bf ``Skill:''} The expectation value of the information gain, 
$\skill$
should  the next subject classified have lens probability 0.5
(\Aref{appendix:swap}),  in bits. Random classifiers have $\skill = 0.0$,
perfect classifiers have $\skill = 1.0$. All agents start with $\skill = 0.0$;
this increases as training subjects are classified, and the agent's estimates
of its confusion matrix elements improve. In stage 1, the mean skill per agent
was 0.04 bits; in  stage 2 it was 0.05.
%
\item{\noindent\bf ``Contribution:''} The integrated skill over a volunteer's test
subject classification history, and represents the total  contribution to the
project that volunteer (see the appendix for more discussion of this
quantity). In stage 1, the mean contribution per agent was 34.9 bits; in 
stage 2 it was 33.5.
%
\item{\noindent\bf ``Information:''} The total information $\information$ generated by
the  agent during the volunteer's classification activity. This quantity
depends on the value of each subject's lens probability when that subject was 
presented to the volunteer (\Aref{appendix:swap}), and so there is an element
of luck involved with this quantity: if you never see a high probability
subject, it's hard to generate a large amount of information. You make your
own luck by classifying more subjects.
%
\end{description}

@@

\Tref{tab:crowd:contributions} shows the total effort, contribution, skill and
information generated in both stage 1 and 2 of the CFHTLS project, with the
total numbers of agents and subjects for comparison. These numbers allow us to
quantify the efficiency of the system. The contribution per classification is
defined in terms of a hypothetical subject with lens probability of 0.5; one
bit of information is needed to update such a subject's lens probability to
either zero or one. This means that a maximally complete classification stage
would yield a total contribution (summed over all agents) equal to the number
of subjects. The ratio of this hypothetical optimum to the actual total
contribution is therefore a measure of the stage's inefficiency. We find our
inefficiency (by dividing column 2 by column 3) to be 33\% and 17\% in stage 1
and 2 respectively. In stage 1, this inefficiency is due to the daily
processing: we were not able to retire subjects fast enough, and so they,
remained in the system, being over-classified. Only 3705745 classifications
were needed to retire all the subjects: the ratio of this to the total number
made is 34\%. (The remaining 1\% is due to not all subjects being classified
to 1 or 0 probability.) At stage 2, we did not retire any subjects at all; the
inefficiency in this case was by design, to give everyone a chance to
appreciate what they had found together!

@@

The total information generated during the survey is harder to interpret,
because it's exactly the high probability candidates that we leave in the
system that give rise to most of the information generated. What we can do is
divide the total information generated by the amount of information it takes
to classify a \sw subject all the way to the detection threshold (lens
probability 0.95), divide by this number, and then multiply by the survey
inefficiency to get a very rough estimate for the effective number of
detections corresponding to the crowd's contribution. With a prior probability
of $2\times10^{-4}$, it takes 12.3 bits to make a detection (see
\Sref{sec:appendix:infogain} in the appendix); the effective
number of detections contributed by the stage 1 and stage 2 volunteers is then
2830 and 25 bits respectively. 
These figures are close to the numbers of detections
given in column 7 of \Tref{tab:crowd:contributions}. 
The uncertainty in the interpretation of the
information generated provides further justification for our focus on the
expected information gain as a measure of volunteer contribution.

@@

The full confusion matrix of the $k^{\rm th}$ volunteer's agent is therefore:
\begin{align}
  \CM^k &= 
  \begin{bmatrix}
    \pr(\saidLENS|\NOT,T_k) & \pr(\saidLENS|\LENS,T_k) \\
    \pr(\saidNOT |\NOT,T_k) & \pr(\saidNOT |\LENS,T_k)
  \end{bmatrix}, \notag \\
        &=
  \begin{bmatrix}
    \CM_{LN} & \CM_{LL} \\
    \CM_{NN} & \CM_{NL}
  \end{bmatrix}^k.
  \label{eq:confmat}
\end{align}
Note that these probabilities are normalized, such that
$\pr(\saidNOT |\NOT) = 1 - \pr(\saidLENS|\NOT)$.

@@

Suppose we have a volunteer who is always right about the true
nature of a training subject. 
Their agent's confusion matrix would be
\be
  \CM^{\rm perfect} = 
  \begin{bmatrix}
    0.0 & 1.0 \\
    1.0 & 0.0
  \end{bmatrix}.
\ee
On being given a fresh subject that actually is a $\LENS$, this hypothetical
volunteer would submit $C = \saidLENS$.  Their agent would then calculate the
posterior probability for the subject being a $LENS$ to be
\begin{align}
  \pr(\LENS|\saidLENS,T_k) &= \frac{1.0 \cdot p_0}
           {\left[ 1.0\cdot p_0 + 0.0\cdot(1 - p_0) \right]}
   &= 1.0,
\end{align}
as we might expect for such a {\it perfect} classifier.  Meanwhile, a
hypothetical volunteer who (for some reason) wilfully always submits the wrong
classification would have an agent with the column-swapped confusion matrix
\be
  \CM^{\rm obtuse} = 
  \begin{bmatrix}
    1.0 & 0.0 \\
    0.0 & 1.0
  \end{bmatrix},
\ee
and would submit $C = \saidNOT$ for this subject. However, such a volunteer
would nevertheless be submitting useful information, since given the above
confusion matrix, their agent would calculate
\begin{align}
  \pr(\LENS|\saidNOT,T_k) &= \frac{1.0 \cdot p_0}
           {\left[ 1.0\cdot p_0 + 0.0\cdot(1 - p_0) \right]}
   &= 1.0.
\end{align}
{\it Obtuse} classifiers turn out to be as helpful as {\it perfect} ones.

@@

\subsection{Agent ``Skill'' and ``Contribution''}
\label{appendix:swap:examples}

% The information likely to be contributed by each agent for a given subject
% can be estimated before the next classification of that subject is made, 
% just from its confusion matrix.
% The Shannon entropy generated by a classifier upon performing a
% classification is
% \be
%   \langle S_k \rangle =  
%    -P_{\rm right}\cdot\log_2{P_{\rm right}} -  P_{\rm wrong}\cdot\log_2P_{\rm wrong}\,,
% \ee
% where $P_{\rm right}$ and $P_{\rm wrong}$ are the averages of the diagonal and
% the off-diagonal elements of the confusion matrix, respectively, and $\langle
% S_k \rangle$ is measured in ``bits.'' These averages represent the probability
% of a classifier to get a classification right or wrong, respectively. We define
% the information contributed by a classifier as
% \begin{equation}
% I_k = 1 - S_k\,.
%   \label{eq:app:info}
% \end{equation}
% \Eref{eq:app:info} gives the required result, that both the hypothetical {\it
% perfect} and {\it obtuse} classifiers contribute 1 bit of information each, per
% classification. Classifiers whose agent's confusion matrix is such that $P_{\rm
% right}=P_{\rm wrong}=0.5$, contribute zero bits of information. Such users
% identify a lens correctly with the same probability as they misclassify a dud
% image to contain a lens, and thus their classification is of no value.

With an agent's confusion matrix in hand we can compute the
\emph{information} generated in any given classification. This will
depend on the confusion matrix elements (\Eref{eq:confmat}) but also on
the probability of the subject being classified containing a lens. The
quantity of interest is the mutual information between the prior and
posterior probabilities given the submitted classification $C$:
\begin{align}
\information =& H(T) - H(T;C) \notag \\ 
             =& -\pr(\LENS) \log_2 \pr(\LENS) \notag \\
              & -\pr(\NOT) \log_2 \pr(\NOT) \notag \\
              & -\pr(\LENS|C) \log_2 \pr(\LENS|C) \notag \\
              & -\pr(\NOT|C) \log_2 \pr(\NOT|C),
\end{align}
where, as above, $C$ can take the values $\saidLENS$ or $\saidNOT$. 
($H$ is the Shannon [cross] entropy function.) 
Substituting for the posterior probabilities using \Eref{eq:app:first} we get
an expression that just depends on the elements of the 
confusion matrix $\CM$ and the pre-classification subject lens
probability $\pr(\LENS) = p$:
\begin{align}
\information =    & -    p \log_2 p - (1-p) \log_2 (1-p) \notag \\
                  & - \log_2 \left[ \CM_{CL} p + \CM_{CN} (1-p) \right] \notag \\
                  & + \frac{1}{p \CM_{CL} + (1-p) \CM_{CN}} \notag \\ 
     \cdot [  p   & \CM_{CL} \left(\log_2 \CM_{CL} + \log_2 p\right) \notag \\
         + (1-p)  & \CM_{CN} \left(\log_2 \CM_{CN} + \log_2 (1-p)\right) ].
  \label{eq:app:infogain}
\end{align}
This expression has many interesting features.  If $p$ is either zero or
one, $\information(C) = 0$  regardless of the value of $C$ or the values
of the confusion matrix elements: if we know the subject's status with
certainty, additional classifications supply no new information. If we
set $p$ to be the prior probability, \Eref{eq:app:infogain} tells us how
much information is generated by classifying it all the way to $p = 1$
(which a perfect classifier, with $\CM_{LL} = \CM_{NN} = 1$, can do in
a single classification). For a prior probability of $2\times 10^{-4}$, 
12.3 bits are generated in such a ``detection.'' 
Conversely, only 0.0003 bits
are generated during the rejection of a subject with the same prior: we
are already fairly sure that each subject does not contain a lens!
Imperfect classifiers (with $\CM_{LL}$ and $\CM_{NN}$ less than 1)
generate less than these maximum amounts of information each
classification; the only classifiers that generate zero information are
those that have $\CM_{LL} = 1 - \CM_{NN}$. We might label such
classifiers as ``random'', since they are as likely to classify a
subject as a $\saidLENS$ no matter the true content of that subject.

\Eref{eq:app:infogain} suggests a useful information theoretical 
definition of the classifier skill  perceived by the agent. At a fixed
value of $p$, we can take the expectation value of the information gain
$\information$ over  the possible classifications that could be made:
\begin{align}
\langle\information\rangle   =& H(T) - \sum_C H(T;C) \pr(C) \notag \\ 
         &   =     p  \left[ \mathcal{S}(\CM_{LL}) + \mathcal{S}(1-\CM_{LL}) \right] \notag \\
         &     +(1-p) \left[ \mathcal{S}(\CM_{NN}) + \mathcal{S}(1-\CM_{NN}) \right] \notag \\
         & - \mathcal{S}\left[ p    \CM_{LL}       + (1-p)(1-\CM_{NN})     \right] \notag \\
         & - \mathcal{S}\left[ p (1-\CM_{LL})      + (1-p)   \CM_{NN}      \right] 
\end{align}
where the function $\mathcal{S}(x) = x \log_2(x)$. If we choose to
evaluate $\langle\information\rangle$ at $p = 0.5$, the result has some
useful properties. While random classifiers presented with  $p = 0.5$
subjects have $\skill = 0.0$  as expected, perfect classifiers appear to
the agents to have  $\skill = 1.0$. This suggests that  $\skill$, the
amount of information we expect to  gain when a classifier is presented
with a 50-50 subject, is a reasonable quantification of 
\emph{normalised skill}. A consequence of this choice is that the
integrated skill (over all agents' histories) should come out to be
equal to the number of subjects in the survey, when the search is
``complete'' (and all subjects are fully classified). Therefore, a
particular agent's integrated skill is a measure of that classifier's
\emph{contribution} to the lens search.

We conservatively initialize both elements of each  agent's confusion
matrix to be 0.5, that of a maximally ambivalent  random classifier, so
that all agents start with zero skill. While  this makes no allowance
for volunteers that actually do have previous experience of what
gravitational lenses look like, we might expect it to help mitigate
against false positives. Anyone who classifies more than one image (by
progressing beyond the tutorial) makes a non-zero information
contribution to the project.

@@

Finally, the confusion matrix obtained from the application of
\Eref{eq:app:fraction} has some inherent noise which reduces as the
number of training subjects classified by the agent's volunteer
increases. For simplicity, the discussion has thus far assumed the case
when the confusion matrix is known perfectly; in practice, we allow for
uncertainty in the agent confusion matrices by averaging over a small
number of samples drawn from Binomial distributions characterised by the
matrix elements $\pr(C_k|\LENS,T_k)$ and  $\pr(C_k|\NOT,T_k)$. The
associated standard deviation in the estimated subject probability
provides an error bar for this quantity.
